---
title: "Lab 6 Comparing two means"
author: "Beau Ayers"
date: 4/26/2021
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

```{r}
# load packages -----------------------------------------------------------

library(tidyverse)


# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish.csv")

# put data in tidy format ------------------------------------------------

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

# do stuff ----------------------------------------------------------------

ttest_results <- t.test(formula = species ~ location, data = fish_long)

ttest_results
```

The mean difference is 1.833333. The 95% confidence interval is -4.587031 to 8.253697.

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

```{r}
ttest_results <- t.test(formula = species ~ location, data = fish_long)

ttest_results
```

There is no evidence that tributries have an effect on the number of species because the p-value is greater than 0.05.

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

```{r}
fish_sum <-
  fish_long %>%
  group_by(location) %>%
  summarize(
    sampl_size = n(),
    mean = mean(species),
    str_dev = sd(species),
    var = var(species),
    sem = sd(species) / sqrt(n()),
    ci_upper = mean + 2 * sem,
    ci_lower = mean - 2 * sem,
  )

fish_sum

ggplot(data = fish_long) +
  geom_jitter(mapping = aes(x = location, y = species)) +
  geom_crossbar(
    data = fish_sum,
    mapping = aes(x = location, y = mean, ymax = ci_upper, ymin = ci_lower),
    color = "cyan3"
  )

fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 8, 
    alpha = 0.5, 
    position = "identity",
    na.rm = TRUE
  ) +
  scale_fill_manual(values = c("darkorange", "darkorchid")) +
  theme_minimal()
```

I had to assume that the p-value was 0.05. That it was a random sample and the distribution was normal

# ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

## Question D

Graph the distribution of body temperatures for each crab type:

```{r}
# Crabs -------------------------------------------------------------------

crabs <- read_csv("chap15q27FiddlerCrabFans.csv")


# Code for crab graphs ----------------------------------------------------

crab_sum <-
  crabs %>%
  group_by(crabType) %>%
  summarize(
    sampl_size = n(),
    mean = mean(bodyTemp),
    str_dev = sd(bodyTemp),
    var = var(bodyTemp),
    sem = sd(bodyTemp) / sqrt(n()),
    ci_upper = mean + 2 * sem,
    ci_lower = mean - 2 * sem,
  )

crab_sum

ggplot(data = crabs) +
  geom_jitter(mapping = aes(x = crabType, y = bodyTemp)) +
  geom_crossbar(
    data = crab_sum,
    mapping = aes(x = crabType, y = mean, ymax = ci_upper, ymin = ci_lower),
    color = "cyan3"
  )

crabs %>% 
  ggplot(aes(x = bodyTemp)) +
  geom_histogram(
    aes(fill = crabType), 
    bins = 8, 
    alpha = 0.5, 
    position = "identity",
    na.rm = TRUE
  ) +
  scale_fill_manual(values = c("darkorange", "darkorchid", "cyan4", "#C24641")) +
  theme_minimal()
```

## Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

```{r}
# ANOVA crabs -------------------------------------------------------------

aov_crabs <-
  aov(bodyTemp ~ crabType, data = crabs)
aov_crabs

summary(aov_crabs)
```

Body temperature does vary among crab types. This is apparent because the F value is greater than 1 and the p-value is far lower than 0.05.

Null: There is no statistical difference in temperature among different crab types

Alternative: There will be a statistical difference in temperature among different crab types
